---
#title: "Subset analyses"
output: md_document
params:
  input_file: NULL
  hlen: NULL
  hlen_exactly: FALSE
  sample_fraction: 1.0
  model: "dt"
  mode: "initial"
  sex: "both"
  id: NULL
  date: "2020-07-08"
  extra_id: NULL
  summary_table_file: NULL
  effect_size_table_file: NULL
  prediction_table_file: NULL
  Hb_cutoff_male: 135
  Hb_cutoff_female: 125
  predictive_variables: NULL
  hyperparameters: "filename"
  cores: 4
  iterations: 2000
  skip_train: FALSE
  create_datasets_bool: TRUE
  donor_specific_file: NULL
  dev: "cairo_pdf"  
---

  
```{r, echo=FALSE, message=FALSE}
message(sprintf(">>>>>>>>>> Decision tree %s %s >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>", params$sex, knitr::opts_chunk$get("dev")))
```

```{r, results="asis", echo=FALSE, message=FALSE}
cat(sprintf("# Decision tree (%s)\n", params$sex))
```


<!-- <h2> Initialization </h2> -->
```{r Setup, setup = TRUE, echo=FALSE, message=FALSE, results="hide"}


suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))
library(rstan)
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(tictoc))
suppressPackageStartupMessages(library(entropy))
suppressPackageStartupMessages(library(brms))
suppressPackageStartupMessages(library(ggmcmc))
suppressPackageStartupMessages(library(ggthemes))
suppressPackageStartupMessages(library(ggridges))
suppressPackageStartupMessages(library(ROCR))
suppressPackageStartupMessages(library(pROC))
suppressPackageStartupMessages(library(furrr))
suppressPackageStartupMessages(library(sn))
suppressPackageStartupMessages(library(tidyverse))

source("common.R")

# This can measure time a chunk took to execute.
# Add chunk option time_it=TRUE to each chunk your want to measure.
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      # record the current time before each chunk
      now <<- lubridate::now()
    } else {
      # calculate the time difference after a chunk
      res <- lubridate::now() - now
      # return a character string to show the time
      msg <- paste("Time for this code chunk to run:", as.numeric(res), units(res))
      message(msg)
      NULL   # Don't return the message so that it won't be printed to the resulting document.
    }
  }
}))


knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, results = "hide", time_it=TRUE)
#knitr::opts_chunk$set(echo=TRUE, message=TRUE)
#options(warn = 1)

if (knitr::opts_chunk$get("dev") == "pdf") {
  knitr::opts_chunk$set(dev="cairo_pdf")
  knitr::opts_current$set(dev="cairo_pdf")
}
message(paste0("Device is ", knitr::opts_chunk$get("dev"), "\n"))
message(paste0("Current device is ", knitr::opts_current$get("dev"), "\n"))
use_pdf <- knitr::opts_chunk$get("dev") %in% c("pdf", "cairo_pdf")

message("Parameters are:")
for (name in names(params)) {
  message(sprintf("%s = ", name), params[[name]])
}

mode <- params$mode



set.seed(123)
number_of_cores <- parallel::detectCores()
if (!is.null(params$cores)) {
  number_of_cores <- min(params$cores, number_of_cores)
  options(mc.cores   = number_of_cores)   # This option is in package parallel
  options(boot.ncpus = number_of_cores)   # For bootstrapping
} else {
  options(mc.cores   = number_of_cores)
  options(boot.ncpus = number_of_cores)
}



datadir = "../data/rdata/"
rawresultdir = "../data/raw_results/"   # I store plot objects and tables in this directory

# Id for a run. For example files in rawresultdir use this id as part of their name. 
# Probably I should add a possibility to download raw results from the web interface.
if (is.null(params$id)) {
  id <- paste(params$date, params$hlen, params$hlen_exactly, params$extra_id, sep="_") 
} else {
  id <- params$id
}

source("helper_functions.R")
source("validate_stan_fit.R")

data <- load_single(params$input_file) # Forgets the name of the object

data <- data %>%
  mutate(days_to_previous_fb = as.double(days_to_previous_fb),
         donor = as.character(donor))   # For some reason this cannot be a factor






# I use this to pass the "smoking", "bmi", "RNF43_mutant", "prs" to the algorithms
if (!is.null(params$donor_specific_file) && (length(params$donor_specific_file) > 0)) {
  donor_specific_variables <- load_single(params$donor_specific_file)
  donor_variables <- donor_descript %>% filter(Variable %in% names(donor_specific_variables)) %>% pull(Variable)
  
  #pretty_donor_variables <- donor_descript$Pretty

  data <- inner_join(data, donor_specific_variables, by="donor")
} else {
  donor_variables = NULL
  donor_specific_variables <- NULL
}




# Make sure the columns in data are in the same order as in the vars list (only needed for stan preprocessing)
variables_in_order <- setdiff(descript$Variable, "donor")
if (!is.null(donor_variables)) {
  message(paste(variables_in_order, collapse=" "))
  message(paste(donor_variables, collapse=" "))
  message(paste(colnames(data), collapse=" "))
  data <- data %>% select(all_of(variables_in_order),
                          all_of(donor_variables), everything())
} else {
  data <- data %>% select(all_of(variables_in_order), everything())
}



# Maybe this is not needed except for linear regression
data <- filter_based_on_number_of_donations(data, params$hlen, params$hlen_exactly)



```



## Data description



### Donation-specific variables

```{r, echo=FALSE, results="markup"}
kable(descript)
```

### Donor-specific variables

```{r, echo=FALSE}
kable(donor_descript %>% filter(Variable %in% names(donor_specific_variables)))
```

```{r Descriptions, results="asis", echo=FALSE}
cat(sprintf("### Summary plots of variables (%s)\n", params$sex))
```


```{r Summary plots, echo = FALSE, results="hide"}
create_summary_plots(data, donor_specific_variables, params$sex, descript, donor_descript)





```{r Partition and enrich, echo=FALSE}
debug <- TRUE

donors <- ndonor(data)
message(sprintf("Number of donors is %i\n", length(donors)))
old_count <- nrow(data); old_count2 <- ndonor(data)

#tmp <- old_split(data)
tmp <- new_split(data, mode)
train_orig <-tmp$train
test <- tmp$test

message(sprintf("Dropped %i / %i donations (%i / %i donors) due to taking training subsample\n", 
              old_count - nrow(train_orig), old_count, old_count2 - ndonor(train_orig), old_count2))
message("Here")
sink(file=stderr(), type="output"); summary(train_orig); sink()
n <- length(unique(train_orig %>% group_by(donor) %>% filter(max(Hb_deferral) == 1) %>% ungroup() %>% pull(donor)))
message(sprintf("%i donors have at least one deferral\n", n))
#message(summary(train_orig))
if (debug) save(train_orig, file="/tmp/pre-enriched.rdata")

# enrich the data so that 50% of the donors have deferral as their last donation attempt
old_count <- nrow(train_orig); old_count2 <- ndonor(train_orig)
#sink(file=stderr(), type="output")
if (old_count2 >= 1000) {
  enriched <- enrich_deferrals_rf(train_orig, 0.5)  # truncates time series
  #sink()
  message(sprintf("Dropped %i / %i donations (%i / %i donors) due to enrichment\n", 
                  old_count - nrow(enriched), old_count, old_count2 - ndonor(enriched), old_count2))
  if (debug) save(enriched, file="/tmp/post-enriched.rdata")
} else {  # Don't enrich if less than 1000 donors
  message(sprintf("Skipped enrichment of train data, since number of donors %i is less than %i", old_count2, 1000))
  enriched <- train_orig
}

n <- length(unique(enriched %>% group_by(donor) %>% filter(max(Hb_deferral) == 1) %>% ungroup() %>% pull(donor)))
message(sprintf("%i donors have at least one deferral\n", n))

# Drop donors whose last ferritin is not measured before the last blood donation after the truncation
if ("FERRITIN_LAST_DATE" %in% names(enriched)) {
  old_count <- nrow(enriched); old_count2 <- ndonor(enriched)
  enriched <- enriched %>% 
    group_by(donor) %>% 
    mutate(last_donation_date = max(dateonly)) %>% 
    ungroup() %>% 
    filter(FERRITIN_LAST_DATE < last_donation_date)  %>% 
    select(-last_donation_date)
  message(sprintf("Dropped %i / %i donations (%i / %i donors) due to FERRITIN_LAST_DATE not being before last blood donation\n", 
              old_count - nrow(enriched), old_count, old_count2 - ndonor(enriched), old_count2))
}
sink(file=stderr(), type="output"); summary(enriched); sink()
if (debug) save(enriched, file="/tmp/enriched.rdata")
#enriched <- enriched %>% select(-FERRITIN_LAST_DATE)
#rm(train_orig)

message("and here")
```






```{r clean up memory, echo = FALSE}
rm(data); invisible(gc())
```





## Results




```{r Decision tree, results="hide"}
id <- sprintf("dt-%s", params$sex)

# Instead of using the real input data, I create random data
n=500   # We create n random rows
if (params$sex == "both") {
  sex2 <- sample(c("male", "female"), n, replace = TRUE)
} else {
  sex2 <- rep(params$sex, n)
}

df <- tibble(id=id,
             model="dt",
             sex = sex2,
             original_value  = rnorm(n, 150, 20),
             predicted_value = rnorm(n, 150, 20),
             original_label  = as.integer(original_value < ifelse(sex=="male", params$Hb_cutoff_male, params$Hb_cutoff_female)),
             predicted_label = as.integer(predicted_value < ifelse(sex=="female", params$Hb_cutoff_male, params$Hb_cutoff_female)),
             score           = rnorm(n))

if (!is.null(params$prediction_table_file)) {
  write_csv(df, params$prediction_table_file)   # Pass the table to the web app as well
}

# Create the plots
results_decision_tree <- gather_results(df, Id=id, Model = "dt", Pretty="Decision tree", Sex=params$sex)

# Show the plots
results_decision_tree$confusion_matrix_plot
results_decision_tree$roc_plot 
results_decision_tree$pr_plot


summary_table <- results_decision_tree$summary

# if (!is.null(params$effect_size_table_file)) {
#   write_csv(rrfFit.varimp, params$effect_size_table_file)   # Pass the table to the web app as well
# }



```







### Summary

```{r Summary}

if (!is.null(params$summary_table_file)) {
  write_csv(summary_table, params$summary_table_file)   # Pass the table to the web app as well
}

```

```{r Show summary table, results="markup"}
cols <- c("Model"="Pretty", "Sex", "MAE (g / L)", "RMSE (g / L)", "MAE (mmol / L)", "RMSE (mmol / L)", "AUROC" = "AUROC value", "AUPR" = "AUPR value", "F1" = "F1 value")
kable(summary_table %>% select(!!!cols), digits=3, format.args = list(digits=3))
```
